---
title: "Arthropod_markdown"
output: github_document
date: '2022-05-05'
---
First retrieve all necessary packages from the library. Note, this (hidden) code also includes a chunk (commented-out) to install Bioc-manager and msa if needed. This chunk is included because their installation directly from R markdown is not possible. 

Packages:
-tidyverse
-knitr
-janitor
-dplyr
-ggplot2
-reshape2
-bold
-msa
-seqRFLP

```{r, setup, include=FALSE}
library(tidyverse)
library(knitr)
library(janitor)
library(dplyr )
library(ggplot2)
library(reshape2)
library(bold)
library(msa)
library(seqRFLP)


#if (!requireNamespace("BiocManager", quietly=TRUE))
#install.packages("BiocManager")
#BiocManager::install("msa")
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install("ggtree")
```

I also read in the "DS-KROK4BDJ.txt" file, which is a tsv downloaded from our paper of choice and (I believe) contains all the data we will need to work with, including sample IDs, geographic information, and sequences. 
```{r,my-chunk, echo=FALSE}
bdj<-read_tsv("DS-KROK4BDJ.txt", show_col_types = FALSE )
column_names <- colnames(bdj)
view(bdj)
```
After reading in the file, I inspect the column dimensions
```{r}
dim(bdj)
```
```{r}
kable(column_names)
```
Let's look at the first 5 rows in this file. It's a large chunk, but very interesting. Scroll side to side to see the whole chunk. 
```{r}
kable(bdj[1:5])
```
From this point, we can pick out columns that we want to work with. For practice, I will select the phylum-subspecies taxID and name columns (14 columns total) and convert them to factor. After this (hidden) code chunk, the "bdj" object (our data) columns with these names will be converted from "numeric" or "character" to "factor". Since they are now factors, we can run other analyses on them and visualize how many of each, etc. 

```{r, chunk, echo=FALSE, include=TRUE}

bdj$phylum_taxID = as.factor(as.numeric(bdj$phylum_taxID))
bdj$class_taxID = as.factor(as.numeric(bdj$class_taxID))
bdj$order_taxID = as.factor(as.numeric(bdj$order_taxID))
bdj$family_taxID = as.factor(as.numeric(bdj$family_taxID))
bdj$genus_taxID = as.factor(as.numeric(bdj$genus_taxID))
bdj$species_taxID = as.factor(as.numeric(bdj$species_taxID))
bdj$subspecies_taxID = as.factor(as.numeric(bdj$subspecies_taxID))
bdj$phylum_name = as.factor(as.character(bdj$phylum_name))
bdj$class_name = as.factor(as.character(bdj$class_name))
bdj$order_name = as.factor(as.character(bdj$order_name))
bdj$family_name = as.factor(as.character(bdj$family_name))
bdj$genus_name = as.factor(as.character(bdj$genus_name))
bdj$species_name = as.factor(as.character(bdj$species_name))
bdj$subspecies_name = as.factor(as.character(bdj$subspecies_name))
```

Next check out the various family names in this file. 

```{r}
unique(bdj$family_name) %>% as.matrix(family_name) -> family_name_view
kable(family_name_view)
```
I'd like to visualize how many samples per family we have. In other words, how many "Therididae"? How many "Anapidae"?

I'll use ggplot to do this. 
```{r}
ggplot(bdj, aes(x=reorder(family_name, family_name, function(x)-length(x)),fill=family_name)) +
  geom_bar() +
  labs(title="Individuals per Family", 
         x="Family Name", y = "Count") +
  theme(axis.text.x = element_text(angle = 90))

```
Do the same thing for genus: 
```{r}

ggplot(bdj, aes(x=reorder(genus_name, genus_name, function(x)-length(x)),fill=genus_name)) +
  geom_bar() +
  labs(title="Individuals per Genus", 
         x="Family Name", y = "Count") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.position="none") 
```
Next read the "bdj' object into a new object (called "phylo"). We can work with the phylo object moving forward. 
```{r}
phylo <- bdj
```
Now we prepare to retrieve the BOLD sequence data. 
```{r}
phylo_bold <- bold_seq(taxon = NULL, ids = phylo$processid, bin = phylo$bin_uri, container = NULL, institutions = NULL, 
                  researchers = NULL, geo = NULL, marker = "COI-5P", response = FALSE)
```

Here we matched the bin_uri information to the BOLD database, retrieved the sequence. And next we checked manually the first few entries against the NCBI database to double check the correct species was identified. We used process_id and bin_uri to define the gene and the species. 

Before we move on, however, we need to check whether there are 124 unique "bin_uri"s in our file. 

```{r}
bdj%>% filter(!duplicated(.[["bin_uri"]])) -> unique_binuri
```
There are only 48 unique bin_uri entries. 

Now we check whether there are 124 unique process ids. 
```{r}
bdj%>% filter(!duplicated(.[["processid"]])) -> unique_processid
```
There are, in fact, 124 unique processids. This is correct and to be expected. 

We also want to know whether there are any NAs in the bin_uri column of phylo. 

```{r}
sum(is.na(phylo$bin_uri))
```
There are 9 NAs in the bin_uri column. This is important to know moving forward because we can expect to return 115 sequences instead of 124. 

Next we make the BOLD output object (phylo_bold) into a dataframe. 

Credit for below code which we used to transform phylo_bold into a dataframe: https://stackoverflow.com/questions/26973029/split-one-row-after-every-3rd-column-and-transport-those-3-columns-as-a-new-row

```{r}
phylo_bold_df<- as.data.frame(matrix(unlist(phylo_bold, use.names = FALSE), ncol = 4, byrow= TRUE))
```

After making the phylo_bold output into a dataframe, we trim out an redundant column and name the three remaining columns "ID", "name", and "sequence". 
```{r}
phylo_bold_trim <- as.tibble(phylo_bold_df) %>%select(1,2,4)
colnames(phylo_bold_trim) = c("ID", "name", "sequence")
```
The final goal for this object is to read into a fasta file. Before we do that, we need to combine the "ID" and "name" columns into one column. After doing this, we check to see what class our object is. We need it to be one dataframe. 

```{r}
phylo_bold_combo<- phylo_bold_trim %>% unite("name", 1:2, na.rm = TRUE, remove = TRUE)
class(phylo_bold_combo)
```
Since we see that this is not one dataframe, but a "tbl_df", "tbl", and "data.frame", we need to change it into a dataframe. After doing this, we make it into a fasta file. 
```{r}
bold_combo_df <- as.data.frame(phylo_bold_combo)

phylo_bold.fasta = dataframe2fas(bold_combo_df, file = "phylo_bold.fasta")

```
We now use the msa package to align the sequences in our fasta file. 
```{r}
mySequencesphylo <- readDNAStringSet("phylo_bold.fasta")
phylo_alignment <- msa(mySequencesphylo)
phylo_alignment
```
phylo_alignment contains aligned sequences. But, moving forward, we will be using the phylo_bold.fasta file and move to python where we will construct the tree. 


